---
title: "On the origin of citations"
author: "sanninnas"
date: "September 15, 2015"
output: 
    html_document:
        toc: true
        number_sections: true
        theme: cerulean
        highlight: espresso
        fig_width: 10
        fig_height: 3
        fig_caption: yes
        self_contained: false
---

# Load the data
## using read.delim

```{r load_data}
counts_raw <- read.delim("data//counts-raw.txt.gz")
counts_norm <- read.delim("data//counts-norm.txt.gz")
```

# Data exploration

What's the distribution of authors in all articles of our data set?
```{r author_histogram, echo=FALSE, fig.cap="Figure 1: Number of Authors per Article", fig.height=8, fig.width=8}
hist(counts_raw$authorsCount, main="Authors per paper", xlab="# authors")
```

Number of facebook shares per article
```{r facebookshare_histogram, echo=FALSE, fig.cap="Figure 2: Number of Facebook Shares per Article"}
hist(counts_raw$facebookShareCount, main="Facebook Popularity", xlab="# Shares")
```
The average number of Facebook shares per paper in the data set is `r mean(counts_raw$facebookShareCount)`

## dplyr
```{r}
library("dplyr")
```

subsetting
```{r}
research <- filter(counts_raw, articleType == "Research Article")
```

```{r}
research_2006 <- filter(research, year ==2006)
nrow(research_2006)
```

```{r}
research_2006_fb <- filter(research, year == 2006, 
                           facebookCommentCount >0)
nrow(research_2006_fb)
```

```{r}
research_2006_fb_tweet_disease <- filter(research, year == 2006, 
                           facebookCommentCount >0 | backtweetsCount >0,
                           grepl("Infectious Diseases", plosSubjectTags))
nrow(research_2006_fb_tweet_disease)
```

```{r}
colnames(research)
```

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
colnames(article_info)
```

```{r}
article_info <- select(research, doi:authorsCount)
colnames(article_info)
```

```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

```{r}
head(select(research, journal))
head(select(research, 3))
```

```{r}
# getting out rows by index
slice(article_info, 1:3)
```

```{r}
# lesson 12 challange 2 'what are people reading but not citing'
low_cite <- research %>% 
    filter(year <= 2008,
                    pdfDownloadsCount > 1000,
                    mendeleyReadersCount > 15,
                    wosCountThru2011 < 10) %>% 
    select(journal, title, year)
nrow(low_cite)
```

### chaining commands with dpylr
pipe character %>%

```{r}
facebook_2006 <- research %>% 
    filter(year == 2006) %>% 
    select(contains("facebook"))
head(facebook_2006)
```

arrange works similar to function order
```{r}
research %>% arrange(desc(authorsCount, wosCountThru2011)) %>%
    select(authorsCount, wosCountThru2011) %>%
    slice(1:10)
```

```{r}
# lesson 13 challenge 1 "Titles of most cited articles"
research %>% arrange(desc(wosCountThru2011)) %>% select(title) %>% slice(1:3) #or head(3)
```

```{r}
# lesson 13 challenge 2 "Lots of authors"
research %>% arrange(desc(authorsCount)) %>% 
    select(authorsCount, title, journal, plosSubjectTags) %>% slice(1:3)
```

### summarizing with dplyr
```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7,
                                yearsSincePublished = weeksSincePublished / 52)
research %>% select(contains("Since")) %>% slice(1:10)
```

using summarize
```{r}
research %>% summarize(plos_mean = mean(plosCommentCount), plos_sd = sd(plosCommentCount), num = n())
```

### using group_by
```{r}
research %>% group_by(journal, year) %>% 
    summarize(tweets_mean = mean(backtweetsCount))
```

```{r}
# lesson 14 challenge
tweets_per_journal <- research %>% 
    group_by(journal) %>%
    summarize(num = n(), tweets_mean = mean(backtweetsCount), sem = sd(backtweetsCount)/sqrt(num))
```

## ggplot2
```{r}
library("ggplot2")
# ggplot2: you have to pass it a dataframe -> mapping each column to a feature of the graph = aesthetic
```

```{r}
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011))
# nothing happens, need to add layers by connecting layers with + sign
p + geom_point()

# or build everything in the same assignment
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011)) +
    geom_point(aes(color = journal)) + # adds on a new variable "journal"" and mapps the color based on that variable
    geom_smooth()
p

p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011,
                                           color = journal)) +
    geom_point()+
    geom_smooth()
p

p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011)) +
    geom_point(aes(size = authorsCount)) # size of the point is porportional to the number of authors
p

p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011)) +
    geom_point(aes(alpha = daysSincePublished)) 
p

## aes for data-dependent 
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011)) +
    geom_point(color = "red") # everything gets assigned red, regardless of the data
p

```

```{r}
# lesson 15 challenge 'Citations vs days since published'
p <- ggplot(data = research, aes(x = daysSincePublished,
                                 y = wosCountThru2011)) +
    geom_point(aes(color = journal), alpha = 0.5) +
    geom_smooth(color = "red")
p
```

### using scales
```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount,
                                 y = wosCountThru2011)) +
    geom_point(aes(color = journal)) +
    geom_smooth()
p + scale_x_log10() + scale_y_log10() # problem with zeros - intercept in the aes call

p <- ggplot(data = research, aes(x = log10(pdfDownloadsCount+1),
                                 y = log10(wosCountThru2011+1))) +
    geom_point(aes(color = journal)) +
    geom_smooth()
p

p <- ggplot(data = research, aes(x = log10(pdfDownloadsCount+1),
                                 y = log10(wosCountThru2011+1))) +
    geom_point(aes(color = journal)) +
    geom_smooth() + 
    scale_x_continuous(breaks = c(1,3), labels = c(10, 1000))
p

p <- ggplot(data = research, aes(x = log10(pdfDownloadsCount+1),
                                 y = log10(wosCountThru2011+1))) +
    geom_point(aes(color = journal)) +
    geom_smooth() + 
    scale_x_continuous(breaks = c(1,3), labels = c(10, 1000)) + 
    scale_y_continuous(breaks = c(1,3), labels = c(10, 1000))
p

# modifying limits
# setting limits: R will give warning about values getting cut off
# scale will geom_smooth calculated with data that are only included after scaling
p <- ggplot(data = research, aes(x = log10(pdfDownloadsCount+1),
                                 y = log10(wosCountThru2011+1))) +
    geom_point(aes(color = journal)) +
    geom_smooth() + 
    scale_x_continuous(breaks = c(1,3), labels = c(10, 1000)) + 
    scale_y_continuous(breaks = c(1,3), labels = c(10, 1000),
                       limits = c(1,3))
p
```

different color options
```{r}
p + scale_color_grey()
p + scale_color_manual(values=c("red", "green", "blue", "orange", "pink", "yellow", "purple"))
```

```{r}
library("RColorBrewer")
display.brewer.all(type = "qual")
```

```{r}
p + scale_color_brewer(palette="Dark2")

# changing labels of the graph
p + scale_color_brewer(palette="Dark2",
                       labels = 1:7, name = "PLOS")
```

```{r}
# lesson 16 challenge
p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
    geom_point(aes(color = journal)) +
    geom_smooth() +
    scale_color_brewer(palette = "Accent")
p  
```

## using facets to make subplots

```{r}
p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
    geom_point(aes(color = journal)) +
    geom_smooth() +
    scale_color_brewer(palette = "Accent")
p + facet_wrap(~journal, ncol = 2)
```

using facet_grid

```{r}
research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags))
p + facet_grid(journal~immuno)
## failed because p was defined without immuno - need to redefine p

p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
    geom_point(aes(color = journal)) +
    geom_smooth() +
    scale_color_brewer(palette = "Accent")
p + facet_grid(journal~immuno)
```


### using different geoms

```{r}
p <- ggplot(data = research, aes(x = journal,
                                 y = sqrt(wosCountThru2011))) +
    geom_boxplot()
p
```

barplots
```{r}
tweets_per_journal <- research %>% 
    group_by(journal) %>%
    summarize(num = n(), tweets_mean = mean(backtweetsCount), sem = sd(backtweetsCount)/sqrt(num))
```

```{r}
tweets_bar <- ggplot(tweets_per_journal, aes(x = journal, y = tweets_mean)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin = tweets_mean - sem, ymax = tweets_mean + sem), width = 0.1) + 
    geom_text(aes(label = num), hjust = 0, vjust = 0)
tweets_bar
```

```{r}
# lesson 18 challenge 1
tweets_per_journal_year <- research %>% 
    group_by(journal, year) %>%
    summarize(num = n(), tweets_mean = mean(backtweetsCount), sem = sd(backtweetsCount)/sqrt(num))

tweets_bar <- ggplot(tweets_per_journal_year, aes(x = journal, y = tweets_mean)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin = tweets_mean - sem, ymax = tweets_mean + sem), width = 0.1) + 
    geom_text(aes(label = num), hjust = 0, vjust = 0)
tweets_bar + facet_wrap(~year, ncol = 2)
```

```{r}
# lesson 18 challenge 2
sd <- ggplot(research, aes(x=log10(wosCountThru2011))) + 
    geom_histogram()
sd
sd <- ggplot(research, aes(x=sqrt(wosCountThru2011))) + 
    geom_histogram()
sd
sd <- ggplot(research, aes(x=log10(wosCountThru2011))) + 
    geom_density()
sd
sd <- ggplot(research, aes(x=sqrt(wosCountThru2011))) + 
    geom_density()
sd
```

### Customizing the plot
```{r}
tweets_bar + labs(title = "Mean tweets per journal per year",
                  x = "Journal", y = "Mean number of tweets") +
    theme_minimal() +
    facet_wrap(~year, ncol = 2)
```

```{r}
# setting themes for the whole chunk's plots
theme_set(theme_bw())
```


## Notes
To add comments in rmarkdown document, use code chunk and set include = FALSE
```{r include=FALSE}
## comments here
```



